\documentclass[10pt]{article}
\include{../../preamble.tex}

\title{Behavioral Pset 1}

\author{Dylan Baker}
\date{October 2024}

\begin{document}
\maketitle

\section{Problem 2}

\subsection{Question 2.1}

We can get a loss aversion parameter by asking about preferences 
over lotteries with possible loss or gain versus sure thing 
degenerate lotteries. Loss aversion would appear as 
accepting a sure thing below the expected value of the lottery
if the lottery has a possible loss. Estimating risk aversion  
could involve a similar set of operations, but with lotteries 
that don't involve a possible loss.

\subsection{Question 2.2}

It depends on what values they're incrementing by flat dollar amounts over. 
If it's incrementing by a flat dollar from $x.99$ to $(x+y).99$
each time, where $x, y \in \mathbb{N}$, 
then this could be taking advantage of left-digit bias.
If they increment by a flat dollar amount that isn't moving 
around a $x.99$ setup, then it's less clear but could 
be related to reducing menu costs or adhering to industry norms.
Alternatively (and maybe less convincingly), you could imagine a strange form of 
left-digit bias committed by the firm 
themselves in which, \emph{the firm}, facing the need to increase prices,
doesn't feel like they're raising 
prices if it goes up by a leading $0$, e.g., a $0.99$ increase, but does 
if it goes up by a leading $1$, e.g., a $1.00$ increase. Thus, it's 
the minimum increase at which the firm feels like they're raising prices
as needed.

\subsection{Question 2.3}

Rational inattention corresponds to the idea that an agent is 
not dedicating attention to a matter, because the attention 
required to assess the matter is not worth the expected 
benefit of the assessment. Irrational inattention is less 
rooted in a probabilistic assessment of expected costs and benefits 
of assessment and more rooted in cognitive biases. How we 
might distinguish between the two seems like it would vary significantly 
by context. In general, one could imagine varying the expected payoff change
associated with 
undergoing the assessment and seeing if the agent's behavior changes (as would be 
expected for rational inattention and not under 
irrational inattention, all else equal) 
or varying something related to the relevant cognitive bias, holding 
the expected payoff change and cost of assessment constant, and seeing if the agent's behavior changes
(as would be expected for irrational inattention and not under rational inattention).

\section{Problem 3}


\subsubsection{Question 3.1.1}

\begin{align}
    u^{\prime}\left(C_t\right)=\delta f^{\prime}\left(K_t\right) u^{\prime}\left(C_{t+1}\right)
\end{align}

\subsubsection{Question 3.1.2}

In this case,

\begin{align}
    u'(C_t) = C_t^{-\sigma}
\end{align}

Thus,

\begin{align}
    u^{\prime}\left(C_t\right)=\delta f^{\prime}\left(K_t\right) u^{\prime}\left(C_{t+1}\right)
\end{align}

becomes

\begin{align}
    &C_t^{-\sigma} = \delta f^{\prime}\left(K_t\right) C_{t+1}^{-\sigma} \\
    \Leftrightarrow &\frac{C_{t+1}}{C_t} = (0.95 f^{\prime}\left(K_t\right))^{\frac{1}{\sigma}} \\ 
    \Leftrightarrow &\frac{C_{t+1} - C_t}{C_t} = (0.95 f^{\prime}\left(K_t\right))^{\frac{1}{\sigma}} - 1
\end{align}

The growth rate of consumption increases with the rate of return on capital investment.
A higher rate of return on capital investment incentivizes delaying consumption.

\subsubsection{Question 3.1.3}

\begin{align}
    \frac{C_{t+1} - C_t}{C_t} = &(0.95 f^{\prime}\left(K_t\right))^{\frac{1}{\sigma}} - 1 \\
    = &\left(\left(0.95\right)\left(1.75\right)\right)^{\frac{1}{2}} - 1 \\
    \approx 0.289
\end{align}

The implied growth rate of consumption is above.

It seems high.

\subsection{Question 3.1.4}

Poverty traps could plausibly explain low investment in 
contexts where the production function is S-shaped and 
individuals get stuck in a low level of capital, but this 
doesn't explain underinvestment amid facing a high rate of return.
``Lumpy'' investments and credit constraints could contribute to explaining
underinvestment amid high returns. This explanation would rely on
an inability to save, invest in more liquid assets to 
build towards the large investment, or borrow/access other financial 
instruments. 



\end{document}